Dissertation Proposal

### 2023-11-08 proposal.tex
all set

The SMoL Tutor is firmly grounded in one technique from cognitive and
educational psychology. The fundamental problem is: how do you fix a
misconception? One approach is to only ``present the right answer'',
for fear that discussing the wrong conception might actually reinforce
it. Instead, there is a body of literature starting from
\cite{posnerTheoryConceptualChange1982} that presents a theory of
conceptual change, at whose heart is the \emph{refutation text}. A
refutation text tackles the misconception directly, discussing and
providing a refutation for the incorrect idea. Several studies
\citep{schroederRefutationTextFacilitates2022} have shown their
effectiveness in multiple other domains.

The SMoL Tutor's content structure is also influenced by work on
\emph{case comparisons} (which draws analogies between examples).
\cite{alfieriLearningCaseComparisons2013a} suggests that asking
(rather than not asking) students to find similarities between cases,
and providing principles \emph{after} the comparisons (rather than
before or not at all), are associated with better learning outcomes.

  [FILL Notional Machine]

  [FILL Language Levels] Prior research argues for the benefits of
\emph{language levels} [CITE]: a langauge has better be taught as a
sequence of languages, where each later language includes more
langauge constructs than the previous one.

>>>>

The fundamental problem about correcting misconceptions is: how do you
fix a misconception? One approach is to only ``present the right
answer'', for fear that discussing the wrong conception might actually
reinforce it. Instead, there is a body of literature starting from
\cite{posnerTheoryConceptualChange1982} that presents a theory of
conceptual change, at whose heart is the \emph{refutation text}. A
refutation text tackles the misconception directly, discussing and
providing a refutation for the incorrect idea. Several studies
\citep{schroederRefutationTextFacilitates2022} have shown their
effectiveness in multiple other domains.

\emph{Case comparisons} draws analogies between examples.
\cite{alfieriLearningCaseComparisons2013a} suggests that asking
(rather than not asking) students to find similarities between cases,
and providing principles \emph{after} the comparisons (rather than
before or not at all), are associated with better learning outcomes.

Several groups of authors have argued for the benefits of visualizing
program executions
\citep{dicksonExperiencesImplementingUtilizing2022,
guoOnlinePythonTutor2013, karnalimUsePythonTutor2017,
napsExploringRoleVisualization2002}. These visualization seems to
increase engadgments and help students understand how programs works.
These work often mention \emph{notional machines}
\citep{duboulayBlackBoxGlass1981}, which refers to the underlying
mental models rather than the visual representations.

Prior research argues for the benefits of \emph{language levels}
\citep{findlerDrSchemeProgrammingEnvironment2002,crestaniExperienceReportGrowing2010}:
a langauge has better be taught as a sequence of languages, where each
later language includes more langauge constructs than the previous
one.
